---
title: "Read data and tidy up data"
author: "Anders Tolver"
date: "21 Mar 2018"
output: 
  html_document:
    toc: TRUE
    keep_md: TRUE
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(hms) ### package for storing times
library(tidyverse)
library(lubridate)
```

# Read in data

Kathrine Bjerre Løppenthin provided the following data files for the project

```{r}
list.files(path = "../data")
```

## Randomisation list

```{r}
data_rand_list <- read_excel(path  = "../data/Intervention og kontrol.xlsx"
                   , sheet = 1)
```

## Baseline data

Excel file containing 9 sheets. 

First (=anamnese) and last (=Søvndagbog) sheets are  empty?

Use instead files `Anamnese_baseline_followup.xls` and `Kopi af Indtast søvndagbog.xls`.

```{r}
data_base_psqi <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 2)
data_base_ess <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 3)
data_base_bristol <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 4)
data_base_fys_akt <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 5)
data_base_dep <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 6)
data_base_fys_funk <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 7)
data_base_eq_5d <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 8)

```

## Follow up

Note that `.` and blanks have been used to indicate  missing values ...

```{r}
data_follow_psqi <- read_excel(path  = "../data/Follow-up data.xls", skip = 1, na = "."
                   , sheet = 1)
data_follow_ess <- read_excel(path  = "../data/Follow-up data.xls", skip = 1, na = "."
                   , sheet = 2)
data_follow_bristol <- read_excel(path  = "../data/Follow-up data.xls", skip = 1, na = "."
                   , sheet = 3)
data_follow_fys_akt <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 4)
data_follow_dep <- read_excel(path  = "../data/Follow-up data.xls", skip = 1, na = "."
                   , sheet = 5)
data_follow_fys_funk <- read_excel(path  = "../data/Follow-up data.xls", skip = 1, na = "."
                   , sheet = 6)
data_follow_eq_5d <- read_excel(path  = "../data/Follow-up data.xls", skip = 1, na = "."
                   , sheet = 7)
```

Remove columns 21+22 for `data_follow_psqi`.

```{r}
data_follow_psqi <- data_follow_psqi[, -c(21, 22)]
```

## 1. Anamnese (Baseline + follow up)

```{r}
data_base_anamnese <- read_excel(path  = "../data/Anamnese_baseline_followup.xls", sheet = 1)
data_follow_anamnese <- read_excel(path  = "../data/Anamnese_baseline_followup.xls", sheet = 2)
```

Renaming variables according to key provided by Kathrine (see file `Variabeloversigt`).

```{r}
data_base_anamnese <- plyr::rename(data_base_anamnese
             , replace = c("age" = "Age", "tend_joi" = "Tend_joi"
                           , "job" = "Job", "smoking" = "Smoking") )
data_follow_anamnese <- plyr::rename(data_follow_anamnese
             , replace = c("tend_joi" = "Tend_joi"
                           , "Koffein" = "koffein") )
names(data_base_anamnese)
names(data_follow_anamnese)
```


## S-dagbog (Baseline + follow up)

# Construction of new variables

## 2. PSQI

Consult `scoringsmanual til PSQI.pdf`

### Check for invalid data entries and correct errors

```{r}
apply(data_base_psqi[, -1], 2, unique)
```

```{r}
apply(data_follow_psqi[, -1], 2, unique)
```

#### Question 1

Change 24:00 to 00:00 and correct some obvious errors

```{r}
data_base_psqi$SP2_Spg_1_BA <- with(data_base_psqi, replace(SP2_Spg_1_BA, SP2_Spg_1_BA == "24.00", "00.00"))
data_base_psqi$SP2_Spg_1_BA
```

Grundet inkonsistent brug af `.` eller `:` har der indsneget sig nogle fejl i indlæsning af data som rettes her (se Excel-ark)

```{r}
data_follow_psqi$SP2_Spg_1_FU <- with(data_follow_psqi, replace(SP2_Spg_1_FU, SP2_Spg_1_FU == "23", "23.00"))
data_follow_psqi$SP2_Spg_1_FU <- with(data_follow_psqi, replace(SP2_Spg_1_FU, SP2_Spg_1_FU == "0.875", "21.00"))
data_follow_psqi$SP2_Spg_1_FU <- with(data_follow_psqi, replace(SP2_Spg_1_FU, SP2_Spg_1_FU == "0.9375", "22.30"))
data_follow_psqi$SP2_Spg_1_FU
```

#### Question 3

```{r}
data_base_psqi$SP2_Spg_3_BA <- with(data_base_psqi, replace(SP2_Spg_3_BA, SP2_Spg_3_BA =="5.30", "05.30"))
data_base_psqi$SP2_Spg_3_BA <- with(data_base_psqi, replace(SP2_Spg_3_BA, SP2_Spg_3_BA =="6.30", "06.30"))
data_base_psqi$SP2_Spg_3_BA <- with(data_base_psqi, replace(SP2_Spg_3_BA, SP2_Spg_3_BA =="7.00", "07.00"))
data_base_psqi$SP2_Spg_3_BA
```

```{r}
data_follow_psqi$SP2_Spg_3_FU <- with(data_follow_psqi, replace(SP2_Spg_3_FU, SP2_Spg_3_FU =="7", "07.00"))
data_follow_psqi$SP2_Spg_3_FU <- with(data_follow_psqi, replace(SP2_Spg_3_FU, SP2_Spg_3_FU =="6.15", "06.15"))
data_follow_psqi$SP2_Spg_3_FU <- with(data_follow_psqi, replace(SP2_Spg_3_FU, SP2_Spg_3_FU =="6.30", "06.30"))
data_follow_psqi$SP2_Spg_3_FU <- with(data_follow_psqi, replace(SP2_Spg_3_FU, SP2_Spg_3_FU =="8.30", "08.30"))
data_follow_psqi$SP2_Spg_3_FU <- with(data_follow_psqi, replace(SP2_Spg_3_FU, SP2_Spg_3_FU =="60.00", "06.00"))
data_follow_psqi$SP2_Spg_3_FU <- with(data_follow_psqi, replace(SP2_Spg_3_FU, SP2_Spg_3_FU =="6", "06.00"))

data_follow_psqi$SP2_Spg_3_FU
```

#### Question 4

```{r}
data_base_psqi$SP2_Spg_4_BA <- with(data_base_psqi, replace(SP2_Spg_4_BA, SP2_Spg_4_BA =="07.00", "7"))
```

#### Change time format of questions 1,3

```{r}
data_base_psqi <- mutate(data_base_psqi, SP2_Spg_1_time_BA = parse_time(gsub("\\.", ":", SP2_Spg_1_BA), format = "%H:%M"), SP2_Spg_3_time_BA = parse_time(gsub("\\.", ":", SP2_Spg_3_BA), format = "%H:%M"), psqi4_in_bed = SP2_Spg_3_time_BA - SP2_Spg_1_time_BA )
data_base_psqi$psqi4_in_bed
```

```{r}
data_follow_psqi <- mutate(data_follow_psqi, SP2_Spg_1_time_FU = parse_time(gsub("\\.", ":", SP2_Spg_1_FU), format = "%H:%M"), SP2_Spg_3_time_FU = parse_time(gsub("\\.", ":", SP2_Spg_3_FU), format = "%H:%M"), psqi4_in_bed = SP2_Spg_3_time_FU - SP2_Spg_1_time_FU )
data_follow_psqi$psqi4_in_bed
```

## PSQI - component scores (Baseline)

### Component 1: Subjective sleep quality

```{r}
data_base_psqi <- mutate(data_base_psqi, psqi1 = SP2_Spg_6_BA)
```

### Component 2: Sleep latency

Check carefully conversion to numerical score ....

```{r}
data_base_psqi$SP2_Spg_2_BA_num <- cut(as.numeric(data_base_psqi$SP2_Spg_2_BA), breaks = c(0, 15, 30, 60, Inf)
    , labels = c(0, 1, 2, 3))
data_base_psqi <- mutate(data_base_psqi, psqi2_tmp = parse_integer(SP2_Spg_2_BA_num) + SP2_Spg_5A_BA)
data_base_psqi <- mutate(data_base_psqi, psqi2 = parse_integer(cut(psqi2_tmp, breaks = c(0, 2, 4, 6, Inf), labels = c(0, 1, 2, 3))))
with(data_base_psqi, table(psqi2, psqi2_tmp))
```

### Component 3: Sleep duration

Check carefully conversion to numerical score ....

```{r}
data_base_psqi <- mutate(data_base_psqi, psqi3 = parse_integer(cut(parse_double(data_base_psqi$SP2_Spg_4_BA), breaks = c(-Inf, 5, 6, 7, Inf)
    , labels = c(3, 2, 1, 0))))
table(data_base_psqi$psqi3, data_base_psqi$SP2_Spg_4_BA )
```

### Component 4: Habitual sleep efficiency

Be very careful with conversion to time differences (bedtime after midnat causes some problems) ...

```{r}
data_base_psqi <- mutate(data_base_psqi, psqi4_in_bed = 3600*24*(psqi4_in_bed < 0) + psqi4_in_bed
       , psqi4_sleep_eff = 100 * 3600 * parse_double(SP2_Spg_4_BA) / parse_double(psqi4_in_bed)
       , psqi4 = parse_integer(cut(psqi4_sleep_eff, breaks = c(0, 65, 75, 85, Inf), right = FALSE, labels = c(3, 2, 1, 0))))

summary(data_base_psqi$psqi4_sleep_eff)
```


### Component 5: Sleep disturbances

Check carefully conversion ...

```{r}
choose_vars <- paste("SP2_Spg_5", LETTERS[2:10], "_BA", sep = "")
choose_vars
data_base_psqi$psqi5_tmp <- apply(select(data_base_psqi, choose_vars), 1, sum)
data_base_psqi <- mutate(data_base_psqi, psqi5 = parse_integer(cut(psqi5_tmp, breaks = c(-Inf, 0, 9, 18, 27, Inf)
                                                     , labels = c(-1, 0, 1, 2, 3))))
table(data_base_psqi$psqi5_tmp, data_base_psqi$psqi5)
```

### Component 6: Use of sleeping medication

```{r}
data_base_psqi <- mutate(data_base_psqi, psqi6 = data_base_psqi$SP2_Spg_7_BA)
```

### Component 7: Daytime dysfunction

```{r}
data_base_psqi <- mutate(data_base_psqi, psqi7_tmp = SP2_Spg_8_BA + SP2_Spg_9_BA
       , psqi7 = parse_integer(cut(psqi7_tmp, breaks = c(-Inf, 0, 2, 4, 6, Inf), labels = 0:4)))
with(data_base_psqi, table(psqi7, psqi7_tmp))
```

### Global PSQI

```{r}
data_base_psqi$psqi_total <- apply(select(data_base_psqi, num_range(prefix = "psqi", range = 1:7)), 1, sum)
```

## PSQI - component scores (Follow up)

### Component 1: Subjective sleep quality

```{r}
data_follow_psqi <- mutate(data_follow_psqi, psqi1 = SP2_Spg_6_FU)
```

### Component 2: Sleep latency

Check carefully conversion to numerical score ....

```{r}
data_follow_psqi$SP2_Spg_2_FU_num <- parse_integer(cut(as.numeric(data_follow_psqi$SP2_Spg_2_FU), breaks = c( 0, 15, 30, 60, Inf)
    , labels = c(0, 1, 2, 3)))
data_follow_psqi <- mutate(data_follow_psqi, psqi2_tmp = SP2_Spg_2_FU_num + SP2_Spg_5A_FU)
data_follow_psqi <- mutate(data_follow_psqi, psqi2 = parse_integer(cut(psqi2_tmp, breaks = c(-Inf, 0, 2, 4, 6)
                                                         , labels = 0:3)))
data_follow_psqi$psqi2_tmp
data_follow_psqi$psqi2

with(data_follow_psqi, table(psqi2, psqi2_tmp))
```

### Component 3: Sleep duration

Check carefully conversion to numerical score ....

```{r}
data_follow_psqi <- mutate(data_follow_psqi, psqi3 = parse_integer(cut(parse_double(data_follow_psqi$SP2_Spg_4_FU), breaks = c(-Inf, 5, 6, 7, Inf)
    , labels = c(3, 2, 1, 0))))
table(data_follow_psqi$psqi3, data_follow_psqi$SP2_Spg_4_FU )
```

### Component 4: Habitual sleep efficiency

Be very careful with conversion to time differences (bedtime after midnat causes some problems) ...

```{r}
data_follow_psqi <- mutate(data_follow_psqi, psqi4_in_bed = 3600*24*(psqi4_in_bed < 0) + psqi4_in_bed
       , psqi4_sleep_eff = 100 * 3600 * parse_double(SP2_Spg_4_FU) / parse_double(psqi4_in_bed)
       , psqi4 = parse_integer(cut(psqi4_sleep_eff, breaks = c(0, 65, 75, 85, Inf), right = FALSE, labels = c(3, 2, 1, 0))))
summary(data_follow_psqi$psqi4_sleep_eff)
```

NB: One subject `08-240713` sleeps more hours than he or she stays in bed resulting in sleep efficiency above 100.

### Component 5: Sleep disturbances

Check carefully conversion ...

```{r}
choose_vars <- paste("SP2_Spg_5", LETTERS[2:10], "_FU", sep = "")
choose_vars
data_follow_psqi$psqi5_tmp <- apply(select(data_follow_psqi, choose_vars), 1, sum)
data_follow_psqi <- mutate(data_follow_psqi, psqi5 = parse_integer(cut(psqi5_tmp, breaks = c(-Inf, 0, 9, 18, 27, Inf)
                                                     , labels = c(-1, 0, 1, 2, 3))))
table(data_follow_psqi$psqi5_tmp, data_follow_psqi$psqi5)
```

### Component 6: Use of sleeping medication

```{r}
data_follow_psqi <- mutate(data_follow_psqi, psqi6 = data_follow_psqi$SP2_Spg_7_FU)
```

### Component 7: Daytime dysfunction

```{r}
data_follow_psqi <- mutate(data_follow_psqi, psqi7_tmp = SP2_Spg_8_FU + SP2_Spg_9_FU
       , psqi7 = parse_integer(cut(psqi7_tmp, breaks = c(-Inf, 0, 2, 4, 6, Inf), labels = 0:4)))
with(data_follow_psqi, table(psqi7, psqi7_tmp))
```

### Global PSQI

```{r}
data_follow_psqi$psqi_total <- apply(select(data_follow_psqi, num_range(prefix = "psqi", range = 1:7)), 1, sum)
```

### Overview of component scores from PSQI


```{r}
select(data_follow_psqi, num_range("psqi", 1:7) )
```

**How to handle missing answers?**

## 3. Epworth Sleepiness Scale (ESS)

### Check raw data values (baseline + follow up)

All items with take data values 0-3

```{r}
apply(data_base_ess[, -1], 2, table)
apply(data_follow_ess[, -1], 2, table)
```

Count number of missing values 

```{r}
apply(data_base_ess[, 1 + 1:8], 2, function(z){sum(is.na(z))})
apply(data_follow_ess[, 1 + 1:8], 2, function(z){sum(is.na(z))})

```

### Compute ESS sum score

```{r}
data_base_ess$ess_total <- apply(data_base_ess[, 1 + 1:8], 1, sum)
data_follow_ess$ess_total <- apply(data_follow_ess[, 1 + 1:8], 1, sum)

summary(data_base_ess$ess_total)
summary(data_follow_ess$ess_total)
```


## 4. Bristol (BRAF)

### Check  raw data values (baseline)

Range of `SP4_Spg_1_BA` should be 0-10

```{r}
table(data_base_bristol[, 1 + 1])
```

Range of `SP4_Spg_2_BA` should be 0-7

```{r}
table(data_base_bristol[, 1 + 2])
```

Range of `SP4_Spg_3_BA` should be 0-2

```{r}
table(data_base_bristol[, 1 + 3])
```

Range of `SP4_Spg_5_BA- SP4_Spg_20_BA` should be 0-3

```{r}
raw_tab <- apply(data_base_bristol[, 1 + 4:20], 2, table)
raw_tab
```

### Check  raw data values (follow up)

Range of `SP4_Spg_1_BA` should be 0-10

```{r}
table(data_follow_bristol[, 1 + 1])
```

Range of `SP4_Spg_2_BA` should be 0-7

```{r}
table(data_follow_bristol[, 1 + 2])
```

Range of `SP4_Spg_3_BA` should be 0-2

```{r}
table(data_follow_bristol[, 1 + 3])
```

Range of `SP4_Spg_5_BA- SP4_Spg_20_BA` should be 0-3

```{r}
raw_tab <- apply(data_follow_bristol[, 1 + 4:20], 2, table)
raw_tab
```

Be careful how to handle NAs ...

Tabulate number of NAs

```{r} 
```

### Compute sum scores from BRAF

```{r}
get_psqi_phys <- function(x){
  m_score <- c(10, 7, 2, 3)
  na_pos <- is.na(x)
  res <- sum(x, na.rm = T)/sum(m_score[!na_pos]) * sum(m_score)
  if((sum(na_pos) > 1)||(any(is.na(x[1:2])))){res <- NA}
  return(res)
}

data_base_bristol$BRAF_phys <- apply(data_base_bristol[2:5], 1, get_psqi_phys)

get_psqi_other <- function(x){
  m_score <- rep(3, length(x))
  na_pos <- is.na(x)
  res <- sum(x, na.rm = T)/sum(m_score[!na_pos]) * sum(m_score)
  if((sum(na_pos) > 1)){res <- NA}
  return(res)
}

data_base_bristol$BRAF_living <- apply(data_base_bristol[, 1 + 5:11], 1, get_psqi_other)
data_base_bristol$BRAF_cog <- apply(data_base_bristol[, 1 + 12:16], 1, get_psqi_other)
data_base_bristol$BRAF_emo <- apply(data_base_bristol[, 1 + 17:20], 1, get_psqi_other)
data_base_bristol$BRAF_total <- apply(dplyr::select(data_base_bristol, dplyr::contains("BRAF")), 1, sum)

data_follow_bristol$BRAF_phys <- apply(data_follow_bristol[2:5], 1, get_psqi_phys)

get_psqi_other <- function(x){
  m_score <- rep(3, length(x))
  na_pos <- is.na(x)
  res <- sum(x, na.rm = T)/sum(m_score[!na_pos]) * sum(m_score)
  if((sum(na_pos) > 1)){res <- NA}
  return(res)
}

data_follow_bristol$BRAF_living <- apply(data_follow_bristol[, 1 + 5:11], 1, get_psqi_other)
data_follow_bristol$BRAF_cog <- apply(data_follow_bristol[, 1 + 12:16], 1, get_psqi_other)
data_follow_bristol$BRAF_emo <- apply(data_follow_bristol[, 1 + 17:20], 1, get_psqi_other)
data_follow_bristol$BRAF_total <- apply(dplyr::select(data_follow_bristol, dplyr::contains("BRAF")), 1, sum)

```

```{r}
summary(data_base_bristol$BRAF_phys)
summary(data_base_bristol$BRAF_living)
summary(data_base_bristol$BRAF_cog)
summary(data_base_bristol$BRAF_emo)
summary(data_base_bristol$BRAF_total)

summary(data_follow_bristol$BRAF_phys)
summary(data_follow_bristol$BRAF_living)
summary(data_follow_bristol$BRAF_cog)
summary(data_follow_bristol$BRAF_emo)
summary(data_follow_bristol$BRAF_total)
```

## 5. Fysisik aktivitetsniveau

Remove all comments from baseline questionnaire (col 2 -)

```{r}
data_base_fys_akt <- data_base_fys_akt[, 1:2]
```

Check raw data values

```{r}
table(data_base_fys_akt$SP5_Spg_1_BA)
table(data_follow_fys_akt$SP5_Spg_1_FU)
```


## 6. Center for Epidemiologic studies depression scale

### Check raw data values (baseline)

Range should be 0-3
```{r}
apply(data_base_dep[, -1], 2, function(z){table(z)})
```

### Check raw data values (follow up)

Range should be 0-3
```{r}
apply(data_follow_dep[, -1], 2, table)
```
### Compute sum scores from CES-D

```{r}
data_base_dep$CES_D_total <- apply(data_base_dep[, -1], 1, sum)
data_follow_dep$CES_D_total <- apply(data_follow_dep[, -1], 1, sum)
```


## 7. Fysisk funktion i det daglige (HAQ)

Ignore scoring for aids/devices so far i.e. consider only items 1-20

```{r}
names(data_base_fys_funk)
```

### Baseline (HAQ)

Check raw data values (allowed range: 0 - 3)

```{r}
apply(data_base_fys_funk[, 1 + 1:20], 2, table)
```

Correct / change invalid entry `12` in item 7

**Confirm method for computing alternative HAQ-DI score**

```{r}
data_base_fys_funk$haq1 <- apply(data_base_fys_funk[, 1 + 1:2], 1, max, na.rm = T)
data_base_fys_funk$haq2 <- apply(data_base_fys_funk[, 1 + 3:4], 1, max, na.rm = T)
data_base_fys_funk$haq3 <- apply(data_base_fys_funk[, 1 + 5:7], 1, max, na.rm = T)
data_base_fys_funk$haq4 <- apply(data_base_fys_funk[, 1 + 8:9], 1, max, na.rm = T)
data_base_fys_funk$haq5 <- apply(data_base_fys_funk[, 1 + 10:12], 1, max, na.rm = T)
data_base_fys_funk$haq6 <- apply(data_base_fys_funk[, 1 + 13:14], 1, max, na.rm = T)
data_base_fys_funk$haq7 <- apply(data_base_fys_funk[, 1 + 15:17], 1, max, na.rm = T)
data_base_fys_funk$haq8 <- apply(data_base_fys_funk[, 1 + 18:20], 1, max, na.rm = T)
data_base_fys_funk$haq_NA <- apply(select(data_base_fys_funk, num_range(prefix = "haq", range = 1:8)), 1, function(z){sum(!is.na(z))})

data_base_fys_funk$haq_total <- apply(select(data_base_fys_funk, num_range(prefix = "haq", range = 1:8)), 1, mean)

data_base_fys_funk$haq_total <- replace(data_base_fys_funk$haq_total, data_base_fys_funk$haq_NA < 6, NA)
```

Check scores for each of eight categories

```{r}
apply(select(data_base_fys_funk, num_range(prefix = "haq", range = 1:8)), 2, table)
```

### Follow up (HAQ)

Check raw data values (allowed range: 0 - 3)

```{r}
apply(data_follow_fys_funk[, 1 + 1:20], 2, table)
```

**Confirm method for computing alternative HAQ-DI score**

Technical note:

With `na.rm = T` then `max` function returns `-Inf` for an empty vector. A user modified version returning `NA` for an empty vector is used here.

```{r}
my_max <- function(x, ...){
  if(all(is.na(x))){
    return(NA)
  }
  else{
    return(max(x, ...))
  }
  
}

data_follow_fys_funk$haq1 <- apply(data_follow_fys_funk[, 1 + 1:2], 1, my_max, na.rm = T)
data_follow_fys_funk$haq2 <- apply(data_follow_fys_funk[, 1 + 3:4], 1, my_max, na.rm = T)
data_follow_fys_funk$haq3 <- apply(data_follow_fys_funk[, 1 + 5:7], 1, my_max, na.rm = T)
data_follow_fys_funk$haq4 <- apply(data_follow_fys_funk[, 1 + 8:9], 1, my_max, na.rm = T)
data_follow_fys_funk$haq5 <- apply(data_follow_fys_funk[, 1 + 10:12], 1, my_max, na.rm = T)
data_follow_fys_funk$haq6 <- apply(data_follow_fys_funk[, 1 + 13:14], 1, my_max, na.rm = T)
data_follow_fys_funk$haq7 <- apply(data_follow_fys_funk[, 1 + 15:17], 1, my_max, na.rm = T)
data_follow_fys_funk$haq8 <- apply(data_follow_fys_funk[, 1 + 18:20], 1, my_max, na.rm = T)
data_follow_fys_funk$haq_NA <- apply(select(data_follow_fys_funk, num_range(prefix = "haq", range = 1:8)), 1, function(z){sum(!is.na(z))})

data_follow_fys_funk$haq_total <- apply(select(data_follow_fys_funk, num_range(prefix = "haq", range = 1:8)), 1, mean)

data_follow_fys_funk$haq_total <- replace(data_follow_fys_funk$haq_total, data_follow_fys_funk$haq_NA < 6, NA)

```

Check scores for each of eight categories

```{r}
apply(select(data_follow_fys_funk, num_range(prefix = "haq", range = 1:8)), 2, table)
```



## 8. EQ-5D Helbredsspørgeskema

Check raw data values (items 1-5)

```{r}
apply(data_base_eq_5d[, -c(1, 7)], 2, table)
apply(data_follow_eq_5d[, -c(1, 7)], 2, table)
```

Check raw data values (items 6)

```{r}
summary(data_base_eq_5d$SP8_Spg_6_BA)
summary(data_follow_eq_5d$SP8_Spg_6_FU)
```

## 9. Søvndagbog

# Export cleaned data

## Extract and combine baseline data

**Overvej hvilken version af join der skal benyttes for at sikre mod slåfejl i id nummer**

```{r}
data_base <- select(data_base_psqi, c(1, num_range(prefix = "psqi", range = 1:7), matches("psqi_total"))) %>% mutate(time = "baseline")

tmp_base <- select(data_base_ess, ID_NUM, ess_total) %>% mutate(time = "baseline")
data_base <- left_join(data_base, tmp_base)

tmp_base <- select(data_base_bristol, ID_NUM, contains("BRAF_")) %>% mutate(time = "baseline")
data_base <- left_join(data_base, tmp_base)

tmp_base <- select(data_base_dep, ID_NUM, CES_D_total) %>% mutate(time = "baseline")
data_base <- left_join(data_base, tmp_base)

data_base
```

