---
title: "Read data and tidy up data"
author: "Anders Tolver"
date: "21 Mar 2018"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
```

# Read in data

Kathrine Bjerre Løppenthin provided the following data files for the project

```{r}
list.files(path = "../data")
```

## Randomisation list

```{r}
data_rand_list <- read_excel(path  = "../data/Intervention og kontrol.xlsx"
                   , sheet = 1)
```

## Baseline data

Excel file containing 9 sheets. 

First (=anamnese) and last (=Søvndagbog) sheets are  empty?

Use instead files `Anamnese_baseline_followup.xls` and `Kopi af Indtast søvndagbog.xls`.

```{r}
data_base_anamnese <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 1)
data_base_psqi <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 2)
data_base_ess <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 3)
data_base_bristol <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 4)
data_base_fys_akt <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 5)
data_base_dep <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 6)
data_base_fys_funk <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 7)
data_base_eq_5d <- read_excel(path  = "../data/Baseline data.xls", skip = 1
                   , sheet = 8)

```

## Follow up

```{r}
data_follow_psqi <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 1)
data_follow_ess <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 2)
data_follow_bristol <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 3)
data_follow_fys_akt <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 4)
data_follow_dep <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 5)
data_follow_fys_funk <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 6)
data_follow_eq5d <- read_excel(path  = "../data/Follow-up data.xls", skip = 1
                   , sheet = 7)
```

## Anamnese (Baseline + follow up)

## S-dagbog (Baseline + follow up)