---
title: "Template - results - numerical outcomes"
author: "Anders Tolver"
date: "2018"
output: 
  html_document:
    toc: TRUE
    keep_md: TRUE
    code_folding: hide
params:
  var_name: "psqi_total"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r init, include=FALSE}
library(lubridate)
library(hms)
library(nlme)
library(DT)
library(tidyverse)
library(forcats)
data_full <- read_delim(file = "../data/data_full.txt", delim = "\t")
```

# Code / comments

This template generates output for numerical/continuous outcome.

The following are computed

## Summary measures

For each combination of assessment time and randomization group we compute

* number of observations (excluding missing data!)
* mean
* standard deviation
* median
* interquantile range (given by 25 % and 75 % quantiles)

For each randomization group summary measures are also computed for changes over time (baseline to follow up).

## Model based estimates

A linear mixed model with fixed effects of time, randomization group and their interaction as well as random effect of subject/ID_NUM is used to extract

* mean estimate for alle combinations of group and time
* within group changes over time
* comparison of within group changes over time across the two groups

The tables produced extract

* estimates
* standard error on estimates
* P-value (for testing if estimate equals to zero)
* lower and upper 95 % confidence interval for estimate

## Source Code

```{r include = F}
sum_na <- function(x){sum(!is.na(x))}
my_funs <- funs(n = sum_na, mean = round(mean(., na.rm = T), digits = 4), sd = round(sd(., na.rm = T), digits = 4), median = median(., na.rm = T), q25 = quantile(., probs = 0.25, na.rm = T), q75 = quantile(., probs = 0.75, na.rm = T))

print_num <- function(varname, df = data_full){
    group_by(df, time, Randomisering) %>% 
    select(time, Randomisering, varname) %>%
    summarize_at(.vars = varname, .funs = my_funs) %>%
    filter(!is.na(Randomisering))
}

print_num_diff <- function(varname, df = data_full){
  select(df, ID_NUM, time, Randomisering, varname) %>%
  spread(key = time, value = varname) %>%
  mutate(change = follow_up - baseline) %>%
  group_by(Randomisering) %>%
  summarise_at(.vars = "change", .funs = my_funs ) %>%
  filter(!is.na(Randomisering)) %>%
  mutate(time = "change")
}

#tmp_diff <- print_num_diff(varname = "psqi_total", df = data_full)
#tmp_full <- print_num(varname = "psqi_total", df = data_full)
#bind_rows(tmp_full, tmp_diff)
```

# Results for variable `r params$var_name`

Here given for variable: `r params$var_name`

### Raw data

```{r}
DT::datatable(bind_rows(print_num(params$var_name), print_num_diff(params$var_name)))
```

Visualization of raw data

```{r}
select(data_full, time, Randomisering, ID_NUM, y = params$var_name) %>%
  filter(!is.na(Randomisering)) %>%
  ggplot() + geom_line(aes(x = time, y = y, color = Randomisering, group = ID_NUM)) + labs(x = "Assessment time", y = params$var_name)
```

### Model based 

```{r}
library(nlme)
form1 <- formula(paste(params$var_name, "~ time * Randomisering"))
mod1 <- lme(form1, random = ~1| ID_NUM, data = data_full, na.action = na.omit)

form2 <- formula(paste(params$var_name, "~ Randomisering + time:Randomisering - 1"))
mod2 <- lme(form2, random = ~1| ID_NUM, data = data_full, na.action = na.omit)

form3 <- formula(paste(params$var_name, "~ time:Randomisering - 1"))
mod3 <- lme(form3, random = ~1| ID_NUM, data = data_full, na.action = na.omit)

tmp_table3 <- round(cbind(summary(mod3)$tTable[1:4, c(1, 2, 5)], intervals(mod3)$fixed[, c(1, 3)]
), digits = 4)[, c(1, 2, 4, 5, 3)]

tmp_table2 <- round(cbind(summary(mod2)$tTable[3:4, c(1, 2, 5)], intervals(mod2)$fixed[3:4, c(1, 3)]
), digits = 4)[, c(1, 2, 4, 5, 3)]

tmp_table1 <- round(c(summary(mod1)$tTable[4, c(1, 2, 5)], intervals(mod1)$fixed[4, c(1, 3)]
), digits = 4)[c(1, 2, 4, 5, 3)]

tmp_tab <- as_tibble(rbind(tmp_table3, tmp_table2, tmp_table1))
tmp_tab <- mutate(tmp_tab, contrast = c("control_base", "control_follow", "intervention_base", "intervention_follow", "control_change", "intervention_change", "group_diff_change"))[c(6, 1:5)]

DT::datatable(tmp_tab)
```

### Model validation

```{r}
form0 <- formula(paste(params$var_name, "~ time * Randomisering + ID_NUM"))
mod_val <- lm(form0, data = data_full)
plot(predict(mod_val), rstandard(mod_val)
     , xlab = "Predicted values", ylab = "Standardized residuals")
abline(h = c(-2, 0, 2))
qqnorm(rstandard(mod_val))
abline(0, 1, col = 2, lty = 2)
```

